DEBS Grand Challenge: A High-Throughput, Scalable Solution for Calculating Frequent Routes and Profitability of New York Taxis

### 2015-05-07 paper/experiments.tex
making the camara copy

Figure \ref{routegraph} shows the process graph used to process frequent route queries. \textit{RouteEventEmitter} reads the data file and emits events with a sequence number. \textit{RouteEmitter} contains a set of threads (equals to the number of \textit{RouteProcessors}) to emit events more efficiently. When processing each line first it reads the data line and creates a \textit{Route} object using longitude and latitude coordinates of pickup and dropoff locations. Then it distributes the events among the \textit{RouteProcessors} using the hash value of the \textit{Route} object. \textit{TopRouteProcessor} aggregates the results and saves them to a file.

@@

Figure \ref{latency_profit} shows the latency variation for both setups. For single JVM, latencies slightly increased with the addition of new instances as expected. But for distributed setups latency linearly increases with the addition of new nodes after 6 \textit{ProfitCalculator} instances. This latency increment is due to longer queues at the \textit{TopProfitProcessor}. In our message ordering process, we can send a message to \textit{TopProfitProcessor} only if all queues are not empty. This allows some processes with less top ten profitable events to delay other process messages. This problem can  possibly fixed by adding a limit to the message queue size. On the other hand in a real time system, this event ordering many not required since event processing happens at real time.

>>>>

Figure \ref{routegraph} shows the process graph used to process frequent routes queries. \textit{RouteEventEmitter} reads the data file and emits events with a sequence number. \textit{RouteEmitter} contains a set of threads (equals to the number of \textit{RouteProcessors}) to emit events more efficiently. When processing each line first it reads the data line and creates a \textit{Route} object using longitude and latitude coordinates of pickup and dropoff locations. Then it distributes the events among the \textit{RouteProcessors} using the hash value of the \textit{Route} object. \textit{TopRouteProcessor} aggregates the results and saves them to a file.

@@

Figure \ref{latency_profit} shows the latency variation for both setups. For single JVM, latencies slightly increased with the addition of new instances as expected. But for distributed setups latency linearly increases with the addition of new nodes after 6 \textit{ProfitCalculator} instances. This latency increment is due to longer queues at the \textit{TopProfitProcessor}. In our message ordering process, we can send a message to \textit{TopProfitProcessor} only if all queues are not empty. This allows some processes with less top ten profitable events to delay other process messages. This problem can  possibly fixed by adding a limit to the message queue size. On the other hand in a real-time system, this event ordering many not required since event processing happens at real-time.

### 2015-05-07 paper/introduction.tex
making the camara copy

If we consider an application that must process the 173 million trips dataset within an year, it would be a trivial task and several simple implementations would satisfy that requirement. However if we consider an application that must process these data within minutes with the added constraint of minimal latency to process each event, we observe following research challenges.
\begin{enumerate}
	\item \textit{Handling thousands of values in real time :}  An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
	\item \textit{Parallel evaluation :}  This application does not inherently support parallel evaluation. Partitioning route sets and cell sets triggers events for subsets, but application requires the events with whole sets. Therefore is it possible to partition data and still generate required results to support parallelism?  
\end{enumerate}

@@

\subsection{Contributions}
We observe following key contributions, which can be used in solving future DEBS challenges as well.
\begin{enumerate}
	\item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity. 
	\item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity. One of the key challenges of DEBS 2014 \cite{jerzak2014debs}  was to find the median of a large time window. We believe our solution is useful in such situations.
	\item We have developed a scalable parallel evaluation technique to process queries efficiently. 
\end{enumerate}

>>>>

If we consider an application that must process the 173 million trips dataset within an year, several simple implementations would satisfy that requirement. However if we consider an application that must process these data within minutes with the added constraint of minimal latency to process each event, we observe following research challenges.
\begin{enumerate}
	\item \textit{Handling thousands of values in real-time :}  An algorithm with O(n) time complexity will not scale for thousands. How can we design an algorithm with O(log n) time complexity?
	\item \textit{Parallel evaluation :}  This application does not inherently support parallel evaluation. Partitioning route sets and cell sets triggers events for subsets, but application requires the events with whole sets. Therefore is it possible to partition data and still generate required results to support parallelism?  
\end{enumerate}

@@

\subsection{Contributions}
We observe following key contributions, which can be used in solving future DEBS grand challenges as well.
\begin{enumerate}
	\item We have designed a set of  generic data structures capable of inserting, updating, removing and retrieving top values (order can be defined using a comparator function) with O(log n) time complexity. 
	\item Using our data structures we have developed an algorithm to find the running median with O(log n) time complexity. One of the key challenges of 2014 problem \cite{jerzak2014debs}  was to find the median of a large time window. We believe our solution is useful in such situations.
	\item We have developed a scalable parallel evaluation technique to process queries efficiently. 
\end{enumerate}

### 2015-05-07 paper/query_evaluation.tex
making the camara copy

In this section, we explain how to evaluate queries with the given data set. For this query evaluation we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.

@@

\begin{algorithm}
\caption{Algorithm to aggregate top ten results from QueryEvaluators and generate global events}
\label{aggregate_algorithm}
\begin{algorithmic}

>>>>

In this section, we explain how to evaluate queries with the given dataset. For this query evaluation, we can use a distributed event stream processing system with the process graph as shown in Figure \ref{twonode_graph} for both queries.

@@

\begin{algorithm}
\caption{Algorithm to aggregate top ten results from \textit{QueryEvaluators} and generate global events}
\label{aggregate_algorithm}
\begin{algorithmic}

### 2015-05-07 paper/query_implementation.tex
making the camara copy

\begin{table}
\centering
\caption{Operations of \textit{NodeList} Data Structure}
\begin{tabular}{|l|l|} \hline
Operation Signature & Time Complexity \\ \hline \hline
add(key : Object, value : NodeValue) & O(log n) \\ \hline 
get(key : Object) : NodeValue & O(1) \\ \hline
remove(key : Object) : NodeValue & O(log n) \\ \hline
decrementPosition(key : Object) & O(log n) \\ \hline
incrementPosition(key : Object) & O(log n) \\ \hline
getTopValues() : List<NodeValue> & O(1) \\ \hline
\end{tabular}
\label{nodelist_api}
\end{table}

@@

One way of implementing this data structure is to use a heap. A heap supports adding and extracting maximum value operations in O(log n) time. However, since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back. On the other hand if we use a doubly linked list to store values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations. Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.  A \textit{Map} which points to values can be used to retrieve values using a key.

@@

\subsection{Frequent Route Query}

The frequent route query entails finding the most frequent routes within the last 30 minutes. Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.  Since queue operations are supported in O(1) time, query evaluations happen in O(log n) time. Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query an event. This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an object called \textit{RouteCount}.

>>>>

\begin{table}
\centering
\caption{Operations of \textit{NodeList} Data Structure}
\begin{tabular}{|l|l|} \hline
Operation Signature & Time Complexity \\ \hline \hline
add(key : Object, value : NodeValue) & O(log n) \\ \hline 
get(key : Object) : NodeValue & O(1) \\ \hline
containsKey(key : Object) : boolean & O(1) \\ \hline
remove(key : Object) : NodeValue & O(log n) \\ \hline
decrementPosition(key : Object) & O(log n) \\ \hline
incrementPosition(key : Object) & O(log n) \\ \hline
getTopValues() : List<NodeValue> & O(1) \\ \hline
\end{tabular}
\label{nodelist_api}
\end{table}

@@

One way of implementing this data structure is to use a heap. A heap supports adding and extracting maximum value operations in O(log n) time. However, since this application requires top 10 values, we need to retrieve maximum value ten times and insert them back. On the other hand if we use a doubly linked list to store values, it will take O(n) time for  \texttt{decrementPostion} and \texttt{incrementPosition} operations. Therefore we can use a doubly linked list to keep first 10 values (this makes time complexity for all operations with this data structure O(1)) and a heap to keep other values to avoid above problems.

@@

\subsection{Frequent Routes Query}

The frequent routes query entails finding the most frequent routes within the last 30 minutes. Algorithm \ref{frequent_route_algorithm} shows the algorithm to process this query using the \textit{NodeList} data structure and a queue to keep the events for the time window.  Since queue operations are supported in O(1) time, query evaluations happen in O(log n) time. Algorithm \ref{frequent_route_algorithm} shows the basic algorithm to evaluate the query for an event. This algorithm assumes there is a global \textit{NodeList} object and route counts are stored in an object called \textit{RouteCount}.

### 2015-05-07 paper/related_work.tex
making the camara copy

\section{Related work}
DEBS Grand Challenge problem naturally fallen into distributed stream processing domain. Many such systems have been developed with different feature sets according to the applications they use. Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries. All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data. Apart from basic communication these frameworks provides other features such as reliable message delivery, state persistence and dynamic membership management.

Several approaches has been developed for DEBS Grand Challenge 2014 \cite{jerzak2014debs}. Predictive Load Management in Smart Grid Environments \cite{mutschler2014predictive} provides a solutions based on EventCore. They have developed an efficient algorithms to find median. Scalable Stateful Stream Processing for Samart Grids \cite{fernandez2014scalable} implemented their solution on SEEP \cite{castro2013integrating} a scalable fault tolerant framework. Solving the Grand Challenge Using an Opensource CEP Engine \cite{perera2014solving} solution is based on WSO2 Complex Event Processor. All these solutions partition the data to scale up the application, provide solutions to predict some missing data and provides algorithmic solutions for efficient query processing. Our solution shares the same ideas along these solutions while addressing the specific issues for 2015 Grand Challenge.

>>>>

\section{Related work}
DEBS Grand Challenge problem naturally fall into the distributed stream processing domain. Many such systems have been developed with different feature sets according to the applications they use. Yahoo S4 \cite{neumeyer2010s4} is used at Yahoo to calculate the click-through rate, Twitter Storm \cite{toshniwal2014storm} is used in Twitter to process tweets, MillWheel \cite{akidau2013millwheel} is used at Google to track trends in web queries. All these systems are based on the Actor model \cite{agha1985actors} and provide a generic API to implement logic to process data. Apart from basic communication these frameworks provide other features such as reliable message delivery, state persistence and dynamic membership management. Spark stream \cite{zaharia2012discretized} provides fault tolerance based on its Resilient Distributed Datasets \cite{zaharia2012resilient}.

Several approaches has been developed for DEBS Grand Challenge 2014 \cite{jerzak2014debs}. Predictive Load Management in Smart Grid Environments \cite{mutschler2014predictive} provides a solution based on EventCore. That involved development of efficient algorithms to find median. Scalable Stateful Stream Processing for Samart Grids \cite{fernandez2014scalable} implemented their solution on SEEP \cite{castro2013integrating} a scalable fault tolerant framework. Solving the Grand Challenge Using an Opensource CEP Engine \cite{perera2014solving} solution is based on WSO2 Complex Event Processor. All these solutions partition the data to scale up the application, provide solutions to predict some missing data and provides algorithmic solutions for efficient query processing. Our solution shares the same ideas along these solutions while addressing the specific issues for 2015 Grand Challenge.

### 2015-05-07 paper/suriarachchi.tex
making the camara copy

\documentclass{acm_proc_article-sp}

@@

\title{A High-throughput, Scalable Solution for the New York Taxi Trips DEBS Grand Challenge Problem}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
Processing complex queries on unbounded event streams in real time, is a challenge for many data processing systems. These systems are expected to process data with reduced latency to generate real time-events, and at high throughput to minimize the required hardware. In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real time with low latency and high throughput. These queries involve processing windows of thousands of records. Firstly, such processing demands efficient data structures and algorithms to minimize the processing overhead. Secondly, the system should partition data to evaluate them in parallel to make it scalable.

In this paper, we present a set of data structures that we designed to evaluate the aforementioned queries with O(log n) time complexity and a data partitioning technique to evaluate them in parallel. We then evaluate our solution on a signal machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN. We are able to process the frequent routes query with the 173 million trips dataset within 5 minutes with less than 4 millisecond latency and the profitable cells query with same dataset within 11 minutes with less than 5 millisecond latency.
\end{abstract}

% A category with the (minimum) three required fields
%A category including the fourth, optional field follows...

\keywords{DEBS Grand Challenge, Distributed Event Processing, Time Complexity} % NOT required for Proceedings

\input{introduction}
\input{query_implementation}
\clearpage
\input{query_evaluation}
\input{experiments}
\input{related_work}
\input{conclustion}
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%Appendix A
% That's all folks!
\end{document}

>>>>

%\documentclass{acm_proc_article-sp}
\documentclass{sig-alternate}

@@

\newfont{\mycrnotice}{ptmr8t at 7pt}
\newfont{\myconfname}{ptmri8t at 7pt}
\let\crnotice\mycrnotice%
\let\confname\myconfname%

\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from Permissions@acm.org.}
\conferenceinfo{DEBS$^,$ 15,} {June 29 - July 3, 2015, OSLO, Norway.}
\copyrightetc{Copyright 2015 ACM \the\acmcopyr}
\crdata{978-1-4503-3286-6/15/06\ ...\$15.00.\\
http://dx.doi.org/10.1145/2611286.2611330.}

\title{DEBS Grand Challenge: A High-Throughput, Scalable Solution for Calculating Frequent Routes and Profitability of New York Taxis}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
Processing complex queries on unbounded event streams in real-time, is a challenge for many data processing systems. These systems are expected to process data with reduced latency to generate real-time events, and at high throughput to minimize the required hardware. In this regard, DEBS Grand Challenge 2015 focuses on evaluating two queries (frequent routes and  profitable cells) in real-time with low latency and high throughput. These queries involve processing windows of thousands of records. Firstly, such processing demands efficient data structures and algorithms to minimize the processing overhead. Secondly, the system should partition data to evaluate them in parallel to make it scalable.

In this paper, we present a set of data structures that we designed to evaluate the aforementioned queries with O(log n) time complexity and a data partitioning technique to evaluate them in parallel. We then evaluate our solution on a single machine as well as in a distributed setting in a commodity cluster of machines over a 1Gbps LAN. We were able to process the frequent routes query with the 173 million trips dataset within 5 minutes with less than 4 millisecond latency and the profitable cells query with same dataset within 11 minutes with less than 5 millisecond latency.
\end{abstract}

% A category with the (minimum) three required fields
\category{C.2.4}{Distributed Systems}{Distributed applications}
%A category including the fourth, optional field follows...
\keywords{Distributed Event Processing, Real-time analytics, Running median, Time Complexity, Event Stream Processing} % NOT required for Proceedings

\input{introduction}
\input{query_implementation}
\input{query_evaluation}
\input{experiments}
\input{related_work}
\input{conclustion}
\newpage
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{paper}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%Appendix A
% That's all folks!
\end{document}
